---
title: 'R tutorial workshop 2 ggplot'
author: "Janice Goh"
date: 'Ran on `r format(Sys.time(), "%Y-%m-%d %H:%M")`'
output:
  html_document:
    code_folding: show
    theme: spacelab
    number_sections: true
    highlight: monochrome
    fig_width: 11
    fig_height: 8.5
    toc: true
    toc_float: true
---
## Workshop objectives

Today, we are going to continue practising data wrangling, as well as learning to do simple plots in ggplot.
Similar to the tidyverse, ggplot was created to make figure making easier compared to baseR 
For today, we are going to study how frequently do we have patients on multiple medications

## Settings 

```{r}
## uncheck if not installed
## install.packages("ggplot2")
setwd("~/Documents/NUS_R_workshops/IntroToR/WS2_ggplot/")
library(ggplot2)
library(tidyverse)
```


# load dataset 

```{r}
data  <- read.csv()

```

## Question 1: How many patients do we have that are on multiple chromic medications? 

Let's generate a list of patients on chronic meds for our team to review 

## generate table of patients on more than 1 chronic medication

```{r}
chronic.meds <- data %>% filter(!medications == "Doxycycline_100mg" & 
                                  !medications == "Paracetamol_500mg") %>%
                  group_by(NRIC) %>%
                  filter() %>% 
                  arrange(names)
View(chronic.meds)
```

# plot this as a histogram 

```{r}
chronic.meds %>% group_by(NRIC, names) %>%
  summarize(n = n()) %>% 
  ggplot(aes(x = n))+
  geom_histogram()+
  theme_bw()
```

## Question 2: which day was the busiest in the month of Oct, and which medication was the most popular?

What graph can help us display such data? 

# generating Oct dataset

```{r}
Oct.data <- data %>% 
            separate(Date, into = c("MM", "DD", "YY"), sep = "/") %>% 
              mutate(DD = as.numeric(DD)) %>%
              filter(MM == "10")
```

# Counting number of line items and patients per day 

```{r}
Oct.summary <- Oct.data %>% group_by(DD) %>%
                summarize("line_items" = ,
                          "patients" = ) %>%
                as.data.frame %>%
                gather(key = "Type", value = "count", -DD)
## checking that we did what we intended to do
Oct.data %>% filter(DD == 10) %>% pull(NRIC) %>% unique %>% length
```

# Plotting our data as a line graph

```{r}
Oct.summary %>%
  ggplot(aes(x =, y = , col = , group = ))+
    geom_point(aes(alpha = 0.2))+
  geom_()+
  theme_bw()
```

# As a bar graph

```{r}
Oct.summary %>%
  ggplot(aes(x = , y = ))+
  geom_bar()+
  #facet_wrap(~Type)+
  theme_bw()
```


# a small subanalysis: is there a correlation between line items and patient count?

```{r}
Oct.summary<- Oct.summary %>%
                reshape2::dcast(DD~Type)
reg<-lm(patients ~ line_items, data = Oct.summary)
Oct.summary %>%
  ggplot(aes(x = line_items, y = patients))+
  geom_point()+
  stat_smooth(method = "lm")+
  geom_text(x=30, y = 45, label = paste0("R^2=", summary(reg)$r.squared))+
  geom_text(x = 30, y = 40, label = paste0("pval=", summary(reg)$coefficients[2,4]))+
  theme_bw()
```




